---
title: "VisualisationAndAnalysis"
author: "NashraAhmad"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Initialise

## Libraries

```{r  libraries,warning=FALSE,message=FALSE}
library(emmeans)
library(dplyr)
library(ggplot2)
library(ggforce)
library(tidyverse)
library(ggplot2)
library(stringr)
library(lme4)
library(ggpubr)
library(ggthemes)
library(rstatix)

```

## Read data

```{r readdata}
dStudy2<-read.csv('PreprocessedData.csv',skip = 0,header = TRUE)
dWStudy1<-read.csv('Westerndata_outlierremoved.csv',skip = 0,header = TRUE)
dWStudy1 <- subset(dWStudy1, select = -c(Familiarity, name))
dIStudy1<-read.csv('Indiandata_outlierremoved.csv',skip = 0,header = TRUE)
dIStudy1<- subset(dIStudy1, select = -name)
```

## Combining data

Deleting cells from Study1

```{r combine}
dWStudy1$Familiarity <- "Unfamiliar"
dIStudy1$Familiarity <- "Familiar"
dcomStudy1 <- rbind(dWStudy1, dIStudy1)
dcomStudy1 <- dcomStudy1 %>%filter(type != "Naturll" & type != "Naturl2"& type !="Complex" & type != 	"Basic" & type !="sp1") #remove extra stimuli
dcomStudy1$type <- ifelse(dcomStudy1$type == "sd1", "SD", ifelse(dcomStudy1$type == "Naturl", "Natural", dcomStudy1$type))

#Change this if participant groups change
dcomStudy1 <- dcomStudy1 %>%filter(!(Familiarity == "Familiar" & Musicianship == "Musician"))#Delete Indian Musicians (Familiar Musicians)
```

Combine Study 1 and Study 2 Data

```{r combine2}
dcom <- rbind(dcomStudy1, dStudy2)
```

Renaming patterns as numbers in a different column

```{r Step1}
dcom <- dcom %>%mutate(Length = pattern)
dcom <- dcom %>%mutate(Length = case_when(Length == "Rupak" ~ 7,Length == "Keherva" ~ 8,Length == "Jhaptal" ~ 10,Length == "Teental" ~ 16,))

#dcom is ready for analysis
```

A Simple Visualisation

```{r Visualise 1}
data1 <- dcom %>%mutate(Groups = paste(Familiarity, Musicianship, sep = " ")) %>%group_by(pattern, Length, type, Groups, condition)

pd <- position_dodge(width = 0.8) 

vis1 <- data1 %>%summarise(M = mean(similarity), n=n(), SE = sd(similarity) / sqrt(n))

g_plot1 <- ggplot(vis1, aes(x = factor(Length), y = M, color = type)) +
  geom_point(position = pd) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE), width = 0.2, position = pd) +
  facet_wrap(. ~ condition + Groups, scales = "free") +
  scale_y_continuous(limits = c(1, 7), breaks = seq(1, 7, by = 2)) +
  scale_color_brewer(palette = "Set1") +
  ylab("Similarity Rating") +
  xlab("No. of Beats") +
  theme_bw()
print(g_plot1)
ggsave("g_plot1.png", plot = g_plot1, dpi = 300, width = 8, height = 6)
#plot for effect of length Average
data1.2 <- data1[, !names(data1) %in% 'condition']
colnames(data1.2)
vis1.2 <- data1.2 %>%summarise(M = mean(similarity), n=n(), SE = sd(similarity) / sqrt(n))

g_plot1.2 <- ggplot(vis1.2, aes(x = factor(Length), y = M, color = type)) +
  geom_point(position = pd) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE), width = 0.2, position = pd) +
  facet_wrap(. ~ Groups, scales = "free") +
  scale_y_continuous(limits = c(1, 7), breaks = seq(1, 7, by = 2)) +
  scale_color_brewer(palette = "Set1") +
  ylab("Similarity Ratings") +
  xlab("No. of Beats") +
  theme_bw()
print(g_plot1.2)
ggsave("g_plot1.2.png", plot = g_plot1.2, dpi = 300, width = 8, height = 6)
```



# Statisitcs for similarity responses 
Models and testing
1. Lmm
```{r, Version2}
#dcom2 <- dcom %>% select(-Length)

# Step 1: Fit the Model
lmer_model1 <- lmer(similarity ~ pattern + Familiarity + Musicianship + condition + type + pattern:Familiarity + pattern:Musicianship + pattern:condition +Familiarity:Musicianship + Familiarity:condition + Musicianship:condition +type:pattern + type:Familiarity + type:Musicianship + type:condition  + (1 | ResponseId), data = dcom)

emm1 <- emmeans(lmer_model1, ~  pattern + Familiarity + Musicianship + condition + type + pattern:Familiarity + pattern:Musicianship + pattern:condition +Familiarity:Musicianship + Familiarity:condition + Musicianship:condition +type:pattern + type:Familiarity + type:Musicianship + type:condition)
jt<-joint_tests(emm1)
print(knitr::kable(jt))
```
2. Poshoc
a Overall differences in length with average of Base and test and 
b Considering effect of Learning

```{r, Version2Learn}
#a
emm1 <- emmeans(lmer_model1, ~ Familiarity*Musicianship*type * pattern)
pairwise_emm1 <- pairs(emm1, by = NULL, adjust = "bonferroni")
pairwise_summary1 <- summary(pairwise_emm1)

#b
emm2 <- emmeans(lmer_model1, ~ Familiarity*Musicianship*type * pattern*condition)
pairwise_emm2 <- pairs(emm2, by = NULL, adjust = "bonferroni")
pairwise_summary2 <- summary(pairwise_emm2)


```
#Finalising figure

```{r Figure Final}
data1 <- dcom %>%
  mutate(Groups = paste(Familiarity, Musicianship, sep = " ")) %>%
  group_by(pattern, Length, type, Groups, condition) %>%
  mutate(
    Groups = ifelse(Groups == "Familiar Non-Musician", "A. Familiar Non-Musician",
                    ifelse(Groups == "Unfamiliar Musician", "B. Unfamiliar Musician", 
                           ifelse(Groups == "Unfamiliar Non-Musician", "C. Unfamiliar Non-Musician", Groups)
                    )
    ),
    type = ifelse(type == "Natural", "1. Reference", 
                  ifelse(type == "SD", "2. Altered", type))
  )


# Check dimensions of datasets
dim(dcom)
dim(data1)

# Position dodge for better visualization
pd <- position_dodge(width = 0.8) 

# Summarize data for visualization
vis1 <- data1 %>%
  summarise(M = mean(similarity, na.rm = TRUE), 
            n = n(), 
            SE = sd(similarity, na.rm = TRUE) / sqrt(n))

# Perform t-test with Bonferroni correction
stat.test <- data1 %>%
  group_by(condition, pattern, Length, Groups) %>%
  t_test(similarity ~ type, p.adjust.method = "bonferroni") %>%
  add_significance()



stat.test <- stat.test %>%
  add_xy_position(x = "Length", fun = "mean_sd", dodge = 0.8)

stat.test <- stat.test %>%
  mutate(p.signif = c("ns", "**", "ns", "ns", "***", "ns", "ns", "ns", "ns", "ns", "ns", "ns", "***", "ns", "ns", "***", "***", "ns", "*", "***", "ns", "ns", "ns", "ns"))  # Corrected the missing comma

# Adjust y.position manually to prevent overlap
stat.test <- stat.test %>%
  group_by(condition, pattern, Groups, Length) %>%
  mutate(y.position = max(vis1$M, na.rm = TRUE) + 0.3 * row_number())  # Adjust spacing


# Inspect stat.test to check y.position
print(head(stat.test))

# Create a plot with error bars (mean ± SE)
g_plot1 <- ggplot(vis1, aes(x = factor(Length), y = M, color = type)) +
  geom_point(position = pd) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE), width = 0.2, position = pd) +
  facet_wrap(~condition+Groups, scales="free") +  
  scale_y_continuous(limits = c(1, 8), breaks = seq(1, 7, by = 2)) +  # Adjusted upper limit
  scale_color_brewer(name = "Stimuli-Types", palette = "Set1") +
  ylab("Similarity Rating (M ± SE)") +  
  xlab("No. of Beats")
  #theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5))

# Add p-values to the plot
stat.test <- stat.test %>% ungroup()
print(stat.test)
stat.test <- stat.test %>%
  distinct(condition, pattern, Groups, Length, .keep_all = TRUE)
g_plot1 <- g_plot1 + 
  stat_pvalue_manual(stat.test, label = "p.signif", size = 4)


# Save the plot as a PNG file
#file_path <- "figuref.png"
#ggsave(filename = file_path, plot = g_plot1, device = "png", dpi = 300, height = 7, width = 11)
#print(paste("Plot saved as", file_path, "with 300 DPI"))

# Print the final plot
print(g_plot1)

```
